### 2018-09-17 notes/cholesky.tex
notes on 2D calculations

\documentclass[a4paper, 11pt]{article}
\usepackage{fullpage, amssymb, amsmath, graphicx}

\newcommand{\mytitle}{Cholesky Decomposition of 2D Kernel}

\begin{document}
\noindent
\large\textbf{\mytitle} \\ \\ Tyler Gordon \\
\normalsize \today 
\ \ \hrulefill
\section{}
In the 1D case, the cholesky decomposition algorithm for a covariance kernel represented by 
a semiseparable matrix with the form
\begin{equation}
K = A + \text{tril}(UV^T) + \text{triu}(VU^T)
\end{equation}
where U and V are both N x J matrices and A is an N x N diagonal matrix is given in in section 
5.1 of Foreman-Mackey et al, 2017 (FM17). 
In the 2D case, the covariance matrix is given by
\begin{equation}
K = A' + \text{tril}(U'V'^T) + \text{triu}(V'U'^T)
\end{equation}
where $U' = U\otimes Q$, $V' = V\otimes I_M$, and $A' = (A - \sigma_n) \otimes Q + \sigma_{nM+i}$ 
where $Q_{i,j}$ specifies the covariance between 
coordinates $x_i$ and $x_j$ in the second dimension, and $I_M$ is the identity matrix with $M$ corresponding 
to the size of $Q$. Because $U'$ and $V'$ for $U$ and $V$ does not alter the semiseperable property of the 
matrices, the decomposition can proceed as in the 1D case with the following alterations to equations 43-45 of 
FM17: 
\begin{align}
\tilde{U}_{(n-1)M+p, (2j-1)M+q} = a_j\cos(d_jt_n) + b_j\sin(d_jt_n)Q_{p,q} \\
\tilde{U}_{(n-1)M+p, (2j)M+q} = a_j\cos(d_jt_n) + b_j\sin(d_jt_n)Q_{p,q} \\
\tilde{V}_{(n-1)M+p, (2j-1)M+q} = a_j\sin(d_jt_n) - b_j\cos(d_jt_n)\delta_{p, q} \\
\tilde{V}_{(n-1)M+p, (2j)M+q} = a_j\sin(d_jt_n) - b_j\cos(d_jt_n)\delta_{p, q} \\
A_{(n-1)M+p, (n-1)M+p} = \sigma_{(n-1)M+p} + Q_{p, p}\sum_{j=1}^J a_j
\end{align}
where $p, q = 1, 2...M$. 
\begin{equation}
	\phi_{(n-1)M+p, (2j-1)M+q} = \phi_{(n-1)M+p, (2j)M+q} = 
	 \begin{cases}
		e^{-c_j(t_n-t_{n-1})} & p = 1 \\
		1 & p > 1
	\end{cases}
\end{equation}
Still with the constraint that
\begin{equation}
	\phi_{1, (2j-1)M+q} = \phi_{1, (2j)M+q} = 0
\end{equation}
The algorithm for computing the cholesky decomposition is unaltered from equation 46 of FM17.

\end{document}

### 2018-09-17 notes/predict.tex
notes on 2D calculations

\documentclass[a4paper, 11pt]{article}
\usepackage{fullpage, amssymb, amsmath, graphicx, bm}

\newcommand{\mytitle}{}

\begin{document}
\noindent
\large\textbf{\mytitle} \\ \\ Tyler Gordon \\
\normalsize \today 
\ \ \hrulefill
\section{Input coordinates in second dimension are same as data coordinates}
In the 1D case, the predictive distribution of a GP at coordinates $\bm{y^*} = (t^*_1, t^*_2... t^*_p)^T$ has a mean given by 
\begin{equation}
	\mu_p^* = \mu_\theta(t_p^*) + K(t_p^*, \bm{t})\bm{z}
\end{equation}
where $\bm{z} = K(\bm{t}, \bm{t})^{-1}[\bm{y} - \bm{\mu_\theta}(\bm{t})]$. In the 2D case for a GP with covariance in the second dimension 
defined the covariance matrix $Q(\bm{u},\bm{u})$, the predictive distribution is given by 
\begin{equation}
	\bm{\mu_p}^* = \bm{\mu_\theta}(t_p^*) + K(t_p^*, \bm{t})X
\end{equation}
where $X_p = Q(\bm{u}, \bm{u})\bm{z}_{pM:(p+1)M}$ is a vector of length M and 
$\bm{z} = K_{2D}^{-1}[\bm{y} - \bm{\mu_\theta}(\bm{t})]$ where $K_{2D}$ is the full 
2D covariance matrix and $\bm{y} = [y(t_1, u_1), y(t_1, u_2), ...y(t_1, u_M), y(t_2, u_1), ...y(t_N, u_M)]^T$ is the dataset on which 
the GP is conditioned. The algorithm for computing the predictive distribution is given by equations B26-B30 in FM17 with the 
substitution of $X$ for $\bm{z}$. The result is a matrix $\mu^* \in \mathbb{Z}^{N x M}$ where $\mu_{n, m}$ is the predicted mean 
at coordinates $(t_n, u_q)$. 
\section{Input coordinates in second dimension are different from data coordinates}
When the input coordinates for the second dimension do not coincide with the data coordinates, the matrix X becomes 
\begin{equation}
	X_p = Q(\bm{u}^*,\bm{u})\bm{z}_{pM:(p+1)M}
\end{equation}
From which we see that $\bm{X}_p$ is just the mean of the predictive distribution for a 1D process with kernel Q at input coordinates $\bm{u}^*$. 
If the covariance in the second dimension is also defined by a celerite kernel, the cost of computing each $\bm{X}_p$ is  
$\mathcal{O}(mM + rR)$ where $M$ is the number of data coordinates in the second dimension, $R$ is the number of 
input coordinates, and $m$ and $r$ are some constants. Therefore the total cost of computing $X$ is $\mathcal{O}(mMP + rRP)$, 
and the total cost of the 2D prediction algorithm is $\mathcal{O}(mMP + rRP + pP + nN) \sim \mathcal{O}((mM+rR+p)P + nN)$
\end{document}

### 2018-09-17 src/gp.jl

@@ -184,7 +184,6 @@ function cholesky_myway!(a_real::Vector{Float64}, c_real::Vector{Float64},
     J_comp = length(a_comp) * M
     J = J_real + 2*J_comp
     
-    eye = zeros(M, M) + I
     
     # initialize the first row of variables (n=1):
     S::Array{Float64, 2} = zeros(J, J)
@@ -205,8 +204,9 @@ function cholesky_myway!(a_real::Vector{Float64}, c_real::Vector{Float64},
         plus = M + q
         U[1, minus] = a_real[l]*Q[p, q]
         U[1, plus] = 0
-        V[1, minus] = eye[p, q]
-        V[1, plus] = 0
+        if p == 1
+            V[1, minus] = 1
+        end
         W[1, minus] = V[1, minus]/D[1]
         W[1, plus] = V[1, plus]/D[1]
     end
@@ -221,8 +221,10 @@ function cholesky_myway!(a_real::Vector{Float64}, c_real::Vector{Float64},
         cd = cos(d_comp[l]*t[1])
         U[1, minus] = (a_comp[l]*cd + b_comp[l]*sd)*Q[p, q]
         U[1, plus] = (a_comp[l]*sd - b_comp[l]*cd)*Q[p, q]
-        V[1, minus] = cd*eye[p, q]
-        V[1, plus] = sd*eye[p, q]
+        if p == q
+            V[1, minus] = cd
+            V[1, plus] = sd
+        end
         W[1, minus] = V[1, minus]/D[1]
         W[1, plus] = V[1, plus]/D[1]
     end
@@ -278,11 +280,6 @@ function cholesky_myway!(a_real::Vector{Float64}, c_real::Vector{Float64},
                 sum_usu += U[i, j]*tmp
                 sum_us[j] += tmp
             end
-            #S[j, j] = (phi[i, j]^2) * (S[j, j] + (W[i-1, j]^2))
-            #sum_usu += 0.5*U[i, j]*S[j, j]*U[i, j]
-            #for k in j:J
-            #    sum_us[j] += U[i, k]*S[j, k]
-            #end
         end
         D[i] = sqrt(diag[r] + Q[p, p]*a_sum - sum_usu)
         for j in 1:J

### 2018-10-15 paper/paper.tex
first commit of paper

%%
%% Beginning of file 'sample62.tex'
%%
%% Modified 2018 January
%%
%% This is a sample manuscript marked up using the
%% AASTeX v6.2 LaTeX 2e macros.
%%
%% AASTeX is now based on Alexey Vikhlinin's emulateapj.cls 
%% (Copyright 2000-2015).  See the classfile for details.

%% AASTeX requires revtex4-1.cls (http://publish.aps.org/revtex4/) and
%% other external packages (latexsym, graphicx, amssymb, longtable, and epsf).
%% All of these external packages should already be present in the modern TeX 
%% distributions.  If not they can also be obtained at www.ctan.org.

%% The first piece of markup in an AASTeX v6.x document is the \documentclass
%% command. LaTeX will ignore any data that comes before this command. The 
%% documentclass can take an optional argument to modify the output style.
%% The command below calls the preprint style  which will produce a tightly 
%% typeset, one-column, single-spaced document.  It is the default and thus
%% does not need to be explicitly stated.
%%
%%
%% using aastex version 6.2
\documentclass{aastex62}

\usepackage{microtype}
\usepackage{url}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{natbib}
\usepackage{multirow}
\bibliographystyle{aasjournal}

%% The default is a single spaced, 10 point font, single spaced article.
%% There are 5 other style options available via an optional argument. They
%% can be envoked like this:
%%
%% \documentclass[argument]{aastex62}
%% 
%% where the layout options are:
%%
%%  twocolumn   : two text columns, 10 point font, single spaced article.
%%                This is the most compact and represent the final published
%%                derived PDF copy of the accepted manuscript from the publisher
%%  manuscript  : one text column, 12 point font, double spaced article.
%%  preprint    : one text column, 12 point font, single spaced article.  
%%  preprint2   : two text columns, 12 point font, single spaced article.
%%  modern      : a stylish, single text column, 12 point font, article with
%% 		  wider left and right margins. This uses the Daniel
%% 		  Foreman-Mackey and David Hogg design.
%%  RNAAS       : Preferred style for Research Notes which are by design 
%%                lacking an abstract and brief. DO NOT use \begin{abstract}
%%                and \end{abstract} with this style.
%%
%% Note that you can submit to the AAS Journals in any of these 6 styles.
%%
%% There are other optional arguments one can envoke to allow other stylistic
%% actions. The available options are:
%%
%%  astrosymb    : Loads Astrosymb font and define \astrocommands. 
%%  tighten      : Makes baselineskip slightly smaller, only works with 
%%                 the twocolumn substyle.
%%  times        : uses times font instead of the default
%%  linenumbers  : turn on lineno package.
%%  trackchanges : required to see the revision mark up and print its output
%%  longauthor   : Do not use the more compressed footnote style (default) for 
%%                 the author/collaboration/affiliations. Instead print all
%%                 affiliation information after each name. Creates a much
%%                 long author list but may be desirable for short author papers
%%
%% these can be used in any combination, e.g.
%%
%% \documentclass[twocolumn,linenumbers,trackchanges]{aastex62}
%%
%% AASTeX v6.* now includes \hyperref support. While we have built in specific
%% defaults into the classfile you can manually override them with the
%% \hypersetup command. For example,
%%
%%\hypersetup{linkcolor=red,citecolor=green,filecolor=cyan,urlcolor=magenta}
%%
%% will change the color of the internal links to red, the links to the
%% bibliography to green, the file links to cyan, and the external links to
%% magenta. Additional information on \hyperref options can be found here:
%% https://www.tug.org/applications/hyperref/manual.html#x1-40003
%%
%% If you want to create your own macros, you can do so
%% using \newcommand. Your macros should appear before
%% the \begin{document} command.
%%
\newcommand{\todo}[3]{{\color{#2}\emph{#1}: #3}}
\newcommand{\durbintodo}[1]{\todo{Durbin}{blue}{#1}}
\newcommand{\gordontodo}[1]{\todo{Gordon}{red}{#1}}

\newcommand{\project}[1]{\textsf{#1}}
\newcommand{\celerite}{\project{celerite}}

%% Tells LaTeX to search for image files in the 
%% current directory as well as in the figures/ folder.
\graphicspath{}

%% Reintroduced the \received and \accepted commands from AASTeX v5.2
\received{}
\revised{}
\accepted{}
%% Command to document which AAS Journal the manuscript was submitted to.
%% Adds "Submitted to " the arguement.
\submitjournal{}

%% Mark up commands to limit the number of authors on the front page.
%% Note that in AASTeX v6.2 a \collaboration call (see below) counts as
%% an author in this case.
%
%\AuthorCollaborationLimit=3
%
%% Will only show Schwarz, Muench and "the AAS Journals Data Scientist 
%% collaboration" on the front page of this example manuscript.
%%
%% Note that all of the author will be shown in the published article.
%% This feature is meant to be used prior to acceptance to make the
%% front end of a long author article more manageable. Please do not use
%% this functionality for manuscripts with less than 20 authors. Conversely,
%% please do use this when the number of authors exceeds 40.
%%
%% Use \allauthors at the manuscript end to show the full author list.
%% This command should only be used with \AuthorCollaborationLimit is used.

%% The following command can be used to set the latex table counters.  It
%% is needed in this document because it uses a mix of latex tabular and
%% AASTeX deluxetables.  In general it should not be needed.
%\setcounter{table}{1}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%% The following section outlines numerous optional output that
%% can be displayed in the front matter or as running meta-data.
%%
%% If you wish, you may supply running head information, although
%% this information may be modified by the editorial offices.
\shorttitle{TBD}
\shortauthors{Gordon et al.}
%%
%% You can add a light gray and diagonal water-mark to the first page 
%% with this command:
% \watermark{text}
%% where "text", e.g. DRAFT, is the text to appear.  If the text is 
%% long you can control the water-mark size with:
%  \setwatermarkfontsize{dimension}
%% where dimension is any recognized LaTeX dimension, e.g. pt, in, etc.
%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% This is the end of the preamble.  Indicate the beginning of the
%% manuscript itself with \begin{document}.

\begin{document}

\title{TBD}

%% LaTeX will automatically break titles if they run longer than
%% one line. However, you may use \\ to force a line break if
%% you desire. In v6.2 you can include a footnote in the title.

%% A significant change from earlier AASTEX versions is in the structure for 
%% calling author and affilations. The change was necessary to implement 
%% autoindexing of affilations which prior was a manual process that could 
%% easily be tedious in large author manuscripts.
%%
%% The \author command is the same as before except it now takes an optional
%% arguement which is the 16 digit ORCID. The syntax is:
%% \author[xxxx-xxxx-xxxx-xxxx]{Author Name}
%%
%% This will hyperlink the author name to the author's ORCID page. Note that
%% during compilation, LaTeX will do some limited checking of the format of
%% the ID to make sure it is valid.
%%
%% Use \affiliation for affiliation information. The old \affil is now aliased
%% to \affiliation. AASTeX v6.2 will automatically index these in the header.
%% When a duplicate is found its index will be the same as its previous entry.
%%
%% Note that \altaffilmark and \altaffiltext have been removed and thus 
%% can not be used to document secondary affiliations. If they are used latex
%% will issue a specific error message and quit. Please use multiple 
%% \affiliation calls for to document more than one affiliation.
%%
%% The new \altaffiliation can be used to indicate some secondary information
%% such as fellowships. This command produces a non-numeric footnote that is
%% set away from the numeric \affiliation footnotes.  NOTE that if an
%% \altaffiliation command is used it must come BEFORE the \affiliation call,
%% right after the \author command, in order to place the footnotes in
%% the proper location.
%%
%% Use \email to set provide email addresses. Each \email will appear on its
%% own line so you can put multiple email address in one \email call. A new
%% \correspondingauthor command is available in V6.2 to identify the
%% corresponding author of the manuscript. It is the author's responsibility
%% to make sure this name is also in the author list.
%%
%% While authors can be grouped inside the same \author and \affiliation
%% commands it is better to have a single author for each. This allows for
%% one to exploit all the new benefits and should make book-keeping easier.
%%
%% If done correctly the peer review system will be able to
%% automatically put the author and affiliation information from the manuscript
%% and save the corresponding author the trouble of entering it by hand.

\correspondingauthor{Tyler A. Gordon}
\email{tagordon@uw.edu}

\author{Tyler A. Gordon}
\affiliation{Department of Astronomy, University of Washington, Box 351580, U.W., Seattle, WA 98195-1580, USA}

\author{Eric Agol}
\affiliation{Department of Astronomy, University of Washington, Box 351580, U.W., Seattle, WA 98195-1580, USA}

%% Note that the \and command from previous versions of AASTeX is now
%% depreciated in this version as it is no longer necessary. AASTeX 
%% automatically takes care of all commas and "and"s between authors names.

%% AASTeX 6.2 has the new \collaboration and \nocollaboration commands to
%% provide the collaboration status of a group of authors. These commands 
%% can be used either before or after the list of corresponding authors. The
%% argument for \collaboration is the collaboration identifier. Authors are
%% encouraged to surround collaboration identifiers with ()s. The 
%% \nocollaboration command takes no argument and exists to indicate that
%% the nearby authors are not part of surrounding collaborations.

%% Mark off the abstract in the ``abstract'' environment. 
\begin{abstract}

Gaussian processes are frequently used to model variability in astrophysical systems. In addition to studying variability in stars, AGN, and X-ray binaries, Gaussian 
process regression can be used to account for correlated noise in light curves of exoplanetary transits and micro-lensing events. One of the chief limitations of this 
technique is the computational time required to compute a Gaussian process model for large data-sets. This also limits prospects for using Gaussian processes with 
multi-dimensional datasets, although sparse-matrix approximations can make GP regression on multi-dimensional datasets feasible for some applications. Here we 
present a method based on the \celerite GP implementation which enables fast computation of multi-dimensional gaussian process noise models for datasets with a 
small second dimension. This method enables the use of GP regression with multi-wavelength transit and micro-lensing light curves.

\end{abstract}

%% Keywords should appear after the \end{abstract} command. 
%% See the online documentation for the full list of available subject
%% keywords and the rules for their use.
\keywords{}

%% From the front matter, we move on to the body of the paper.
%% Sections are demarcated by \section and \subsection, respectively.
%% Observe the use of the LaTeX \label
%% command after the \subsection to give a symbolic KEY to the
%% subsection for cross-referencing in a \ref command.
%% You can use LaTeX's \ref and \label commands to keep track of
%% cross-references to sections, equations, tables, and figures.
%% That way, if you change the order of any elements, LaTeX will
%% automatically renumber them.
%%
%% We recommend that authors also use the natbib \citep
%% and \citet commands to identify citations.  The citations are
%% tied to the reference list via symbolic KEYs. The KEY corresponds
%% to the KEY in the \bibitem in the reference list below.

\section{Introduction} \label{sec:intro}

%% If you wish to include an acknowledgments section in your paper,
%% separate it off from the body of the text using the \acknowledgments
%% command.
\acknowledgments

%% To help institutions obtain information on the effectiveness of their 
%% telescopes the AAS Journals has created a group of keywords for telescope 
%% facilities.
%
%% Following the acknowledgments section, use the following syntax and the
%% \facility{} or \facilities{} macros to list the keywords of facilities used 
%% in the research for the paper.  Each keyword is check against the master 
%% list during copy editing.  Individual instruments can be provided in 
%% parentheses, after the keyword, but they are not verified.

%\vspace{5mm}
%\facilities{}

%% Similar to \facility{}, there is the optional \software command to allow 
%% authors a place to specify which programs were used during the creation of 
%% the manusscript. Authors should list each code and include either a
%% citation or url to the code inside ()s when available.

%\software{}

%% Appendix material should be preceded with a single \appendix command.
%% There should be a \section command for each appendix. Mark appendix
%% subsections with the same markup you use in the main body of the paper.

%% Each Appendix (indicated with \section) will be lettered A, B, C, etc.
%% The equation counter will reset when it encounters the \appendix
%% command and will number appendix equations (A1), (A2), etc. The
%% Figure and Table counter will not reset.

%\appendix

%% The reference list follows the main body and any appendices.
%% Use LaTeX's thebibliography environment to mark up your reference list.
%% Note \begin{thebibliography} is followed by an empty set of
%% curly braces.  If you forget this, LaTeX will generate the error
%% "Perhaps a missing \item?".
%%
%% thebibliography produces citations in the text using \bibitem-\cite
%% cross-referencing. Each reference is preceded by a
%% \bibitem command that defines in curly braces the KEY that corresponds
%% to the KEY in the \cite commands (see the first section above).
%% Make sure that you provide a unique KEY for every \bibitem or else the
%% paper will not LaTeX. The square brackets should contain
%% the citation text that LaTeX will insert in
%% place of the \cite commands.

%% We have used macros to produce journal name abbreviations.
%% \aastex provides a number of these for the more frequently-cited journals.
%% See the Author Guide for a list of them.

%% Note that the style of the \bibitem labels (in []) is slightly
%% different from previous examples.  The natbib system solves a host
%% of citation expression problems, but it is necessary to clearly
%% delimit the year from the author name used in the citation.
%% See the natbib documentation for more details and options.

\begin{thebibliography}{}

% \bibitem[Astropy Collaboration et al.(2013)]{2013A&A...558A..33A} Astropy Collaboration, Robitaille, T.~P., Tollerud, E.~J., et al.\ 2013, \aap, 558, A33

\end{thebibliography}

%% This command is needed to show the entire author+affilation list when
%% the collaboration and author truncation commands are used.  It has to
%% go at the end of the manuscript.
%\allauthors

%% Include this line if you are using the \added, \replaced, \deleted
%% commands to see a summary list of all changes at the end of the article.
%\listofchanges

\end{document}

% End of file `sample62.tex'.

### 2018-10-16 paper/paper.tex
tectonic doesn't like microtype package
fixed todo
outline paper

@@
 
-\usepackage{microtype}
 \usepackage{url}
@@
 \newcommand{\todo}[3]{{\color{#2}\emph{#1}: #3}}
-\newcommand{\durbintodo}[1]{\todo{Durbin}{blue}{#1}}
+\newcommand{\agoltodo}[1]{\todo{Agol}{blue}{#1}}
 \newcommand{\gordontodo}[1]{\todo{Gordon}{red}{#1}}
@@
 \newcommand{\project}[1]{\textsf{#1}}
-\newcommand{\celerite}{\project{celerite}}
+\newcommand{\celerite}{\project{celerite }}
 
@@
 present a method based on the \celerite GP implementation which enables fast computation of multi-dimensional gaussian process noise models for datasets with a 
-small second dimension. This method enables the use of GP regression with multi-wavelength transit and micro-lensing light curves.
+small second dimension. This method enables the use of GP regression with multi-wavelength transit and micro-lensing light curves, and may have applications to radial 
+velocity and direct imaging data as well.
 
@@
 %% keywords and the rules for their use.
-\keywords{}
+%\keywords{}
 
@@
 %% to the KEY in the \bibitem in the reference list below. 
+\tableofcontents
 
-\section{Introduction} \label{sec:intro}
+\section{introduction}\label{sec:intro}
+
+\section{gaussian processes and the celerite model}
+
+\section{2d gaussian processes}
+	\subsection{kronecker structure of the 2d covariance matrix}
+	\subsection{performance}
+
+\section{example: exomoons}
+
+\section{discussion}
+	\subsection{other applications}
+
+\section{conclusion}
 
@@
 
-%\appendix
+\appendix
 

### 2018-10-17 paper/paper.tex
working on the paper...

@@
 %% using aastex version 6.2
-\documentclass{aastex62}
+\documentclass[modern]{aastex62}
 
@@
 %%
+
+% todos
 \newcommand{\todo}[3]{{\color{#2}\emph{#1}: #3}}
@@
 
+% projects
 \newcommand{\project}[1]{\textsf{#1}}
 \newcommand{\celerite}{\project{celerite }}
+
+% useful commands
+\newcommand{\bvec}[1]{{\ensuremath{\boldsymbol{#1}}}}
+\newcommand{\T}{\ensuremath{\mathrm{T}}}
+\newcommand{\expandvec}[2]{\left(\begin{array}{ccccc} #1\quad && \cdots\quad && #2 \end{array}\right)}
 
@@
 %% this information may be modified by the editorial offices.
-\shorttitle{TBD}
+\shorttitle{Celerite2D}
 \shortauthors{Gordon et al.}
@@
 
-\title{TBD}
+\title{}
 
@@
 \begin{abstract}
-
-Gaussian processes are frequently used to model variability in astrophysical systems. In addition to studying variability in stars, AGN, and X-ray binaries, Gaussian 
+Gaussian processes are frequently used to model variability in astrophysical time-series. In addition to studying variability in stars and AGN, Gaussian 
 process regression can be used to account for correlated noise in light curves of exoplanetary transits and micro-lensing events. One of the chief limitations of this 
@@
 multi-dimensional datasets, although sparse-matrix approximations can make GP regression on multi-dimensional datasets feasible for some applications. Here we 
-present a method based on the \celerite GP implementation which enables fast computation of multi-dimensional gaussian process noise models for datasets with a 
-small second dimension. This method enables the use of GP regression with multi-wavelength transit and micro-lensing light curves, and may have applications to radial 
-velocity and direct imaging data as well.
+present a method based on the \celerite GP implementation which enables fast computation of two-dimensional gaussian process noise models for datasets with a 
+small second dimension, such as multi-band photometry. This method enables the use of GP regression with multi-wavelength transit and micro-lensing light curves, 
+and may have applications to radial velocity and direct imaging data as well.
 
@@
 \section{introduction}\label{sec:intro}
+	Gaussian processes can be used to model correlated variability in astrophysical time-series. They have been used to study stellar variability, to analyze AGN 
+	variability, and to model correlated noise in photometry and radial velocity observations of exoplanet host stars \gordontodo{more applications outside of exoplanets? 
+	X-ray binaries, modeling dust distributions, etc...}.
+
+	Gaussian process noise models are most 
+	advantageous in cases where correlated noise dominates over white noise. This is already the relevant case for state-of-the-art radial velocity observations, but transit 
+	photometry is not far behind. High-precision photometry from future space-based observatories such as James Webb will enable detection of ever more shallow 
+	\gordontodo{shallower?}
+	transit signals, so long as they can be distinguished from correlated variability. \gordontodo{This is where we come in...}
+
+	\gordontodo{Are GPs applicable to instrumental noise as well? Seems best to account for instrumental effects with models that account for known factors since 
+	they might not be well-described by a gaussian process, right?}
+
+	\gordontodo{What about other methods of accounting for astrophysical noise? i.e. cosine filtering, polynomials, etc.}
 
 \section{gaussian processes and the celerite model}
+	A Gaussian process is a stochastic noise model which consists of a mean:
+	\begin{equation} 
+		\bvec{\mu_\theta}(\bvec{x}) = \expandvec{\mu_\bvec{\theta}(x_1)}{\mu_\bvec{\theta}(x_N)}
+	\end{equation}
+	and covariance defined the kernel function $k_\bvec{\alpha}(x_n, x_m)$. The mean and kernel functions depend on the parameters 
+	$\bvec{\theta}$ and $\bvec{\alpha}$ respectively, which are referred to as ``hyperparameters" of the GP. The kernel function defines a covariance matrix 
+	$\left[K\right]_{n, m} = k_\alpha(x_n, x_m)$. This model has a likelihood function given by
+	\begin{equation}
+		\ln\ \mathcal{L(\bvec{\theta}, \bvec{\alpha})} = \ln\ p(\bvec{y}|X, \bvec{\theta}, \bvec{\alpha}) = 
+			-\frac{1}{2}\bvec{r_\theta}^\T K_\bvec{\alpha}^{-1}\bvec{r_\theta} 
+			-\frac{1}{2}\ln\ \mathrm{det}(K_\bvec{\alpha}) - \frac{N}{2}\ln(2\pi)
+	\end{equation}
+	\gordontodo{I'm using $X$ here because DFM does, but wouldn't $\bvec{x}$ be more clear?}
+
+	where 
+	\begin{equation}
+		\bvec{y} = \expandvec{y_1}{y_N}
+	\end{equation}
+	are the data, $N$ is the number of datapoints, and
+	\begin{equation}
+		\bvec{r_\theta} = \bvec{y}-\bvec{\mu_\theta}(\bvec{x})
+	\end{equation}
+	By maximizing the likelihood (or, in practice, minimizing the negative of the log-likelihood), one can obtain an estimate of the hyperparameters $\bvec{\theta}$ and 
+	$\bvec{\alpha}$. Furthermore, Markov Chain Monte Carlo (MCMC) methods can be used to sample the likelihood in hyperparameter-space. \gordontodo{expand 
+	on this.}
+	\subsection{The \celerite model}
+		The primary limitation of Gaussian process noise models is that they are computationally expensive, with computation of the likelihood function scaling as 
+		$\mathcal{O}(N^3)$ due to the matrix inversion. This is especially problematic for use cases such as MCMC which require calling the likelihood function many 
+		times. This has prevented the adoption of Gaussian process noise models for datasets larger than $N \approx 10^3$, except in cases where approximate 
+		methods suffice. The \celerite method overcomes this issue by reducing the time to compute the likelihood function to $\mathcal{O}(NJ^2)$ for one 
+		class of likelihood functions; where J is the number of \celerite terms making up the kernel function of the GP. This class of kernel functions are given by 
+		\begin{equation}
+			k_\alpha(x_n, x_m) = \sigma_n^2 \delta_{n, m} + \sum_{j=1}^J a_j e^{-c_j\tau_{n, m}}
+		\end{equation}
+		where $\bvec{\alpha} = (a_1...a_j, c_1...c_j)$ and $\sigma_n^2$ is the variance of the white noise. The coefficients $a$ and $c$ may be complex.
+
+		\gordontodo{Add discussion of the versatility of this kernel, and it's connection to stochastically driven harmonic oscillators.}
 

### 2018-10-18 paper/paper.tex
some math

@@
 		\begin{equation}
-			k_\alpha(x_n, x_m) = \sigma_n^2 \delta_{n, m} + \sum_{j=1}^J a_j e^{-c_j\tau_{n, m}}
+			k_\alpha(x_n, x_m) = \sigma_n^2 \delta_{nm} + \sum_{j=1}^J a_j e^{-c_j\tau_{nm}}
 		\end{equation}
-		where $\bvec{\alpha} = (a_1...a_j, c_1...c_j)$ and $\sigma_n^2$ is the variance of the white noise. The coefficients $a$ and $c$ may be complex.
+		where $\bvec{\alpha} = (a_1...a_j, c_1...c_j)$, $\sigma_n^2$ is the variance of the white noise, and $\tau_{nm} = |x_n-x_m|$. 
+		The coefficients $a$ and $c$ may be complex.
 
@@
 
+		\gordontodo{Add discussion of semiseperability --- the reader should understand that the limitation of \celerite is that only kernels of this form can be 
+		represented by a semi-seperable matrix, which is what makes moving to two dimensions slightly more complicated than ordinary.}
+
 \section{2d gaussian processes}
 	\subsection{kronecker structure of the 2d covariance matrix}
+	In two dimensions the kernel function depends on two coordinates. Whereas above we used $\bvec{x}$ to refer to the single coordinate, we will now use
+	\begin{equation}
+		\bvec{t} = \expandvec{t_1}{t_N}
+	\end{equation}
+	to refer to the first coordinate, and
+	\begin{equation}
+		\bvec{u} = \expandvec{u_1}{u_M}
+	\end{equation}
+	to refer to the second coordinate. In most cases, we will demand that $M << N$. We then redefine $\bvec{x}$ to be the list of points
+	\begin{equation}
+		\bvec{x} = \expandvec{(t_1, u_1), (t_1, u_2), \cdots (t_1, u_M)}{(t_N, u_M)}
+	\end{equation}
+	We can also write this as
+	\begin{equation}
+		x_i = (t_{\lfloor(i-1)/M\rfloor + 1}, u_{i - \lfloor(i-1)/M\rfloor M})
+	\end{equation} 
+	The covariance between $x_i$ and $x_j$ is then given by
+	\begin{equation}
+		k_{\bvec{\alpha}\bvec{\beta}}(x_i, x_j) = \sigma_i^2\delta_{ij} + q_\bvec{\alpha}(\tau_{ij})\circ r_\bvec{\beta}(\nu_{ij})
+	\end{equation}
+	where $\tau_{ij} = |t_{\lfloor(i-1)/M\rfloor + 1}-t_{\lfloor(j-1)/M\rfloor + 1}|$, 
+	$\nu_{ij} = |u_{i - \lfloor(i-1)/M\rfloor M}-u_{j - \lfloor(j-1)/M\rfloor M}|$; 
+	$q$ and $r$ are one-dimensional kernel functions which depend on the hyperparameters $\bvec{\alpha}$ and 
+	$\bvec{\beta}$ respectively; 
+	and $\circ$ is some binary operation. 
+	If $Q_\bvec{\alpha}$ and $R_\bvec{\alpha}$ are the covariance matrices corresponding to 
+	$q_\bvec{\alpha}$ and $r_\bvec{\alpha}$ such that
+	\begin{equation}
+		\left[Q_\bvec{\alpha}\right]_{ij} = q_\bvec{\alpha}(\tau_{ij})\ \ \ \mathrm{and}\ \ \ \left[R_\bvec{\beta}\right]_{ij} = r_\bvec{\beta}(\tau_{ij})
+	\end{equation} 
+	then, using $q_{ij}$ and $r_{ij}$ as shorthand for the entries in $Q_\bvec{\alpha}$ and $R_\bvec{\beta}$, $K_\bvec{\alpha\beta}$ can be represented as: 
+	\begin{equation}
+		K_\bvec{\alpha\beta} = \Sigma + 
+			\begin{bmatrix}
+				r_{11}\circ Q_\bvec{\alpha} & ... & r_{1N}\circ Q_\bvec{\alpha} \\
+				\vdots & \ddots & \\
+				r_{N1}\circ Q_\bvec{\alpha} & ... & r_{NN}\circ Q_\bvec{\alpha}
+			\end{bmatrix}
+	\end{equation}
+	where $[\Sigma]_{ij} = \sigma_i^2\delta_{ij}$ and
+	\begin{equation}
+		r_{pq}\circ Q_\bvec{\alpha} = 
+			\begin{bmatrix}
+				r_{pq}\circ q_{11} & ... & r_{pq}\circ q_{1M} \\
+				\vdots & \ddots & \\
+				r_{pq}\circ q_{M1} & ... & r_{pq}\circ q_{MM}
+			\end{bmatrix}
+	\end{equation}
+	If the operator $\circ$ represents multiplication, we have 
+	\begin{equation}
+		K_\bvec{\alpha\beta} = \Sigma + R_\bvec{\beta} \otimes  Q_\bvec{\alpha}
+	\end{equation}
+	where $\otimes$ is the Kronecker product. 
 	\subsection{performance}

### 2018-10-18 paper/paper.tex
more math

@@
 	\gordontodo{I'm using $X$ here because DFM does, but wouldn't $\bvec{x}$ be more clear?}
-
 	where 
@@
 		methods suffice. The \celerite method overcomes this issue by reducing the time to compute the likelihood function to $\mathcal{O}(NJ^2)$ for one 
-		class of likelihood functions; where J is the number of \celerite terms making up the kernel function of the GP. This class of kernel functions are given by 
+		class of kernel functions; where J is the number of \celerite terms making up the kernel function of the GP. This class of kernel functions are given by 
 		\begin{equation}
-			k_\alpha(x_n, x_m) = \sigma_n^2 \delta_{nm} + \sum_{j=1}^J a_j e^{-c_j\tau_{nm}}
+			k_\alpha(t_n, t_m) = \sigma_n^2 \delta_{nm} + \sum_{j=1}^J a_j e^{-c_j\tau_{nm}}
 		\end{equation}
-		where $\bvec{\alpha} = (a_1...a_j, c_1...c_j)$, $\sigma_n^2$ is the variance of the white noise, and $\tau_{nm} = |x_n-x_m|$. 
+		where $\bvec{\alpha} = (a_1...a_j, c_1...c_j)$, $\sigma_n^2$ is the variance of the white noise, and $\tau_{nm} = |t_n-t_m|$. 
 		The coefficients $a$ and $c$ may be complex.
+
+		For kernel functions of this form, the covariance matrix can be written as a semi-separable matrix of rank $2J$:
+		\begin{equation}
+			K = A + \mathrm{tril}(UV^\T) + \mathrm{triu}(VU^\T)
+		\end{equation}
+		where $U$ and $V$ have entries given by: 
+		\begin{eqnarray}
+			U_{n, 2j-1} &=& a_je^{-c_jt_n}\cos(d_jt_n) + b_je^{-c_jt_n}\sin(d_jt_n) \nonumber \\
+			U_{n, 2j} &=& a_je^{-c_jt_n}\sin(d_jt_n) - b_je^{-c_jt_n}\cos(d_jt_n) \nonumber \\
+			V_{m, 2j-1} &=& e^{c_jt_m}\cos(d_jt_m) \nonumber \\
+			V_{m, 2j} &=& e^{c_jt_m}\sin(d_jt_m)
+		\end{eqnarray}
+		and A is a diagonal matrix:
+		\begin{equation}
+			A_{n,n} = \sigma_n^2 + \sum_{j=1}^Ja_j
+		\end{equation}
+		Cholesky decomposition can be performed on this matrix in $\mathcal{O}(NJ^2)$ time, as described in the appendix.
 
@@
 	\subsection{kronecker structure of the 2d covariance matrix}
-	In two dimensions the kernel function depends on two coordinates. Whereas above we used $\bvec{x}$ to refer to the single coordinate, we will now use
+	In two dimensions the kernel function depends on two coordinates. We will continue to use
 	\begin{equation}
@@
 	\end{equation}
-	to refer to the first coordinate, and
+	to refer to the first coordinate, and introduce
 	\begin{equation}
@@
 	\end{equation}
-	to refer to the second coordinate. In most cases, we will demand that $M << N$. We then redefine $\bvec{x}$ to be the list of points
+	to refer to the second coordinate. In most cases, we will demand that $M << N$. We then define $\bvec{x}$ to be the list of points
 	\begin{equation}
@@
 			\begin{bmatrix}
-				r_{11}\circ Q_\bvec{\alpha} & ... & r_{1N}\circ Q_\bvec{\alpha} \\
+				q_{11}\circ R_\bvec{\beta} & ... & q_{1N}\circ R_\bvec{\beta} \\
 				\vdots & \ddots & \\
-				r_{N1}\circ Q_\bvec{\alpha} & ... & r_{NN}\circ Q_\bvec{\alpha}
+				q_{N1}\circ R_\bvec{\beta} & ... & q_{NN}\circ R_\bvec{\beta}
 			\end{bmatrix}
@@
 	\begin{equation}
-		r_{pq}\circ Q_\bvec{\alpha} = 
+		q_{pq}\circ R_\bvec{\alpha} = 
 			\begin{bmatrix}
-				r_{pq}\circ q_{11} & ... & r_{pq}\circ q_{1M} \\
+				q_{pq}\circ r_{11} & ... & q_{pq}\circ r_{1M} \\
 				\vdots & \ddots & \\
-				r_{pq}\circ q_{M1} & ... & r_{pq}\circ q_{MM}
+				q_{pq}\circ r_{M1} & ... & q_{pq}\circ r_{MM}
 			\end{bmatrix}
@@
 	\begin{equation}
-		K_\bvec{\alpha\beta} = \Sigma + R_\bvec{\beta} \otimes  Q_\bvec{\alpha}
+		K_\bvec{\alpha\beta} = \Sigma + Q_\bvec{\alpha} \otimes  R_\bvec{\beta}
 	\end{equation}
-	where $\otimes$ is the Kronecker product. 
+	where $\otimes$ is the Kronecker product. If the covariance in $\bvec{t}$ is described by a \celerite kernel can write 
+	\begin{eqnarray}
+		K &=& \Sigma + \left[\mathrm{tril}(UV^\T) + \mathrm{triu}(VU^\T)\right]\otimes R \\
+		&=& \Sigma + \mathrm{tril}\left[ (U\otimes R)(V\otimes I_M)^\T\right] + \mathrm{triu}\left[ (V\otimes I_M)(U\otimes R)^\T\right]
+	\end{eqnarray}
+	This covariance has the same semi-separable form as in the one-dimensional case, but now $K$ is an $(NM\times NM)$ matrix of rank $2JM$. 
+	This means that we can apply the same Cholesky decomposition algorithm (described in the appendix) as in the 1d case to compute the GP in 
+	$\mathcal{O}(NJ^2M^3)$.
+
 	\subsection{performance}
+	Computation of a \celerite GP takes $\mathcal{O}(NJ)$ where $N$ is the length of the coordinate vector $\bvec{x}$ and $J$ is the number of 
+	\celerite terms making up the kernel function.
 

### 2018-10-19 paper/paper.tex
new sections
filled in some comments

@@
 \section{introduction}\label{sec:intro}
-	Gaussian processes can be used to model correlated variability in astrophysical time-series. They have been used to study stellar variability, to analyze AGN 
-	variability, and to model correlated noise in photometry and radial velocity observations of exoplanet host stars \gordontodo{more applications outside of exoplanets? 
-	X-ray binaries, modeling dust distributions, etc...}.
+	Gaussian processes have been used extensively in astrophysics to model stochastic variability across space and time. 
+	They have been used to study variable stars and AGN, to model the extragalactic distribution of dust, and to model correlated noise in 
+	photometric and radial velocity observations of exoplanet host stars.	\gordontodo{X-ray binaries?}
 
-	Gaussian process noise models are most 
-	advantageous in cases where correlated noise dominates over white noise. This is already the relevant case for state-of-the-art radial velocity observations, but transit 
-	photometry is not far behind. High-precision photometry from future space-based observatories such as James Webb will enable detection of ever more shallow 
-	\gordontodo{shallower?}
-	transit signals, so long as they can be distinguished from correlated variability. \gordontodo{This is where we come in...}
+	\gordontodo{The following paragraph might make more sense in the exomoons section.}
+	Gaussian process noise models are most advantageous when correlated noise dominates over white noise. 
+	This is already the case for state-of-the-art radial velocity observations, and transit 
+	photometry is not far behind. Future space-based observatories such as James Webb 
+	will enable high-precision photometry in which stellar variability is the main barrier to detecting ever more shallow transit signals. 
+	Stellar variability which obscures transit signal can be a result of spot and faculae crossings, asteroseismic oscillations, or variable 
+	granulation on the star's surface. GPs are able to \gordontodo{something about how a GP model is non-parametric and the advantages thereof? Also agnostic 
+	to the source of variability (seems related)}
+
+	\gordontodo{This is where we come in...}
 
@@
 
-	\gordontodo{What about other methods of accounting for astrophysical noise? i.e. cosine filtering, polynomials, etc.}
+	\gordontodo{What about other methods of accounting for astrophysical noise? i.e. cosine filtering, polynomials, etc. --- also related to GP as non-parametric 
+	noise model}
 
@@
 	This means that we can apply the same Cholesky decomposition algorithm (described in the appendix) as in the 1d case to compute the GP in 
-	$\mathcal{O}(NJ^2M^3)$.
+	$\mathcal{O}(NJ^2M^3)$. While our implementation only handles the Kronecker product case, 
+	his result holds in the case that $\circ$ is an operation other than multiplication, so long as $\circ$ has the property
+	\begin{equation}
+		(a\circ B)(c\circ D) = (ac)\circ(BD)
+	\end{equation}
+	for scalars $a$ and $b$ and matrices $B$ and $D$. 
+	%where $a\circ B$ is the matrix 
+	%\begin{equation}
+	%	a\circ B = \begin{bmatrix}
+	%		a\circ b_{11} & a\circ b_{12} & \cdots \\
+	%		a\circ b_{21} &  a\circ b_{22} & \\
+	%		\vdots & & \ddots & 
+	%	\end{bmatrix}
+	%\end{equation}
+	\gordontodo{See appendix for proof}
 
-	\subsection{performance}
-	Computation of a \celerite GP takes $\mathcal{O}(NJ)$ where $N$ is the length of the coordinate vector $\bvec{x}$ and $J$ is the number of 
-	\celerite terms making up the kernel function.
+	\subsection{sampling the distribution}
+	A gaussian process can be used to simulate correlated noise. If $L$ is the lower triangular factor from the Cholesky decomposition of the 
+	$(N\times N)$ covariance matrix $K$, and $\bvec{z}$ is a vector of length $N$ containing random numbers drawn for a standard normal distribution, i.e.
+	\begin{equation}
+		\bvec{z} \sim \mathcal{N}(1, 0)
+	\end{equation}
+	then $\bvec{x} = \bvec{\mu_x} + L\bvec{z}$ is a vector with entries correlated by $K$ and with mean $\bvec{\mu_x}$, i.e.
+	\begin{equation}
+		\bvec{x} \sim \mathcal{N}(\bvec{\mu_x}, K)
+	\end{equation}
+	\gordontodo{fix this so that $\mu$ depends on $\theta$ and is a function of $\bvec{t}$ or something like that. Compuational cost should be $\mathcal{O}(NJM^2)$}
+
+	\subsection{interpolation and extrapolation}	
+	\gordontodo{Bit more complicated --- if both dimensions are celerite processes then the scaling is better than in the case where $R$ is not celerite. You have notes on this.}
 
 \section{example: exomoons}
+	\gordontodo{What if I used Kepler-1625 as an example? Could simulate WFC3 GRISM data and show that moon could be validated with Celerite2D?}
 

### 2018-10-19 paper/paper.tex
reworked some stuff

@@
 \section{introduction}\label{sec:intro}
-	Gaussian processes have been used extensively in astrophysics to model stochastic variability across space and time. 
-	They have been used to study variable stars and AGN, to model the extragalactic distribution of dust, and to model correlated noise in 
-	photometric and radial velocity observations of exoplanet host stars.	\gordontodo{X-ray binaries?}
+	A Gaussian process is a generalization of the Gaussian distribution to functions \citep{Rasmussen:2006}. They have been used 
+	extensively in astrophysics as a model of stochastic variability across both space and time to study variable stars, AGN, the extragalactic 
+	dust distribution, and to model correlated noise in photometric and radial velocity measurements of exoplanet hosts. \gordontodo{X-ray binaries?}
 
-	\gordontodo{The following paragraph might make more sense in the exomoons section.}
-	Gaussian process noise models are most advantageous when correlated noise dominates over white noise. 
-	This is already the case for state-of-the-art radial velocity observations, and transit 
-	photometry is not far behind. Future space-based observatories such as James Webb 
-	will enable high-precision photometry in which stellar variability is the main barrier to detecting ever more shallow transit signals. 
-	Stellar variability which obscures transit signal can be a result of spot and faculae crossings, asteroseismic oscillations, or variable 
-	granulation on the star's surface. GPs are able to \gordontodo{something about how a GP model is non-parametric and the advantages thereof? Also agnostic 
-	to the source of variability (seems related)}
+	\gordontodo{something about how a GP model is non-parametric and the advantages thereof? Also agnostic 
+	to the source of variability (seems related) --- essentially: what makes a GP noise model desirable over alternatives?}
+
+	Unfortunately, most GP operations are computationally expensive. The common tasks of computing a likelihood function, sampling from a GP, and 
+	interpolating or smoothing data all run in $\mathcal{O}(N^3)$ time for $N$ data points \gordontodo{make sure this is true for interpolation}. This is 
+	especially problematic for use cases that require iterative calls to the likelihood function, as is the case when minimizing the likelihood or performing 
+	Markov Chain Monte Carlo simulations. This has prevented the adoption of GP methods for datasets larger than $N \sim 10^3$.
+
+	The issue of computational expense can be partially overcome either by approximating a GP (often by exploiting the properties of sparse 
+	matrices), or by restricting the user to a subset of GPs for which the computation can be sped up. One member of the latter class is the 
+	\celerite method, which increases the speed of most common GP computations to $\mathcal{O}(NJ^2)$ where $J$ is typically small compared 
+	to $N$. \celerite achieves this increased speed by limiting the user to GPs with a specific functional form of the covariance. Additionally, \celerite 
+	only allows for one-dimensional GPs.
 
 	\gordontodo{This is where we come in...}
-
-	\gordontodo{Are GPs applicable to instrumental noise as well? Seems best to account for instrumental effects with models that account for known factors since 
-	they might not be well-described by a gaussian process, right?}
-
-	\gordontodo{What about other methods of accounting for astrophysical noise? i.e. cosine filtering, polynomials, etc. --- also related to GP as non-parametric 
-	noise model}
 
@@
 		times. This has prevented the adoption of Gaussian process noise models for datasets larger than $N \approx 10^3$, except in cases where approximate 
-		methods suffice. The \celerite method overcomes this issue by reducing the time to compute the likelihood function to $\mathcal{O}(NJ^2)$ for one 
+		methods suffice.
+
+		The \celerite method overcomes this issue by reducing the time to compute the likelihood function to $\mathcal{O}(NJ^2)$ for one 
 		class of kernel functions; where J is the number of \celerite terms making up the kernel function of the GP. This class of kernel functions are given by 
@@
 	\end{equation}
-	then $\bvec{x} = \bvec{\mu_x} + L\bvec{z}$ is a vector with entries correlated by $K$ and with mean $\bvec{\mu_x}$, i.e.
+	then $\bvec{x} = \bvec{\mu_\bvec{\theta}(\bvec{x})} + L\bvec{z}$ is a vector with entries correlated by $K$ and with mean $\bvec{\mu_\theta}$, i.e.
 	\begin{equation}
-		\bvec{x} \sim \mathcal{N}(\bvec{\mu_x}, K)
+		\bvec{x} \sim \mathcal{N}(\bvec{\mu_\theta}, K)
 	\end{equation}
@@
 	\gordontodo{What if I used Kepler-1625 as an example? Could simulate WFC3 GRISM data and show that moon could be validated with Celerite2D?}
+	Gaussian process noise models are most advantageous when correlated noise dominates over white noise. 
+	This is already the case for state-of-the-art radial velocity observations, and transit 
+	photometry is not far behind. Future space-based observatories such as James Webb 
+	will enable high-precision photometry in which stellar variability is the main barrier to detecting ever more shallow transit signals. 
+	Stellar variability which obscures transit signal can be a result of spot and faculae crossings, asteroseismic oscillations, or variable 
+	granulation on the star's surface.
+
+	\gordontodo{Are GPs applicable to instrumental noise as well? Seems best to account for instrumental effects with models that account for known factors since 
+	they might not be well-described by a gaussian process, right?}
+
+	\gordontodo{What about other methods of accounting for astrophysical noise? i.e. cosine filtering, polynomials, etc. --- also related to GP as non-parametric 
+	noise model}
 

### 2018-10-31 paper/paper.tex
added notebook to make plots for NSF proposal

@@
 	\celerite method, which increases the speed of most common GP computations to $\mathcal{O}(NJ^2)$ where $J$ is typically small compared 
-	to $N$. \celerite achieves this increased speed by limiting the user to GPs with a specific functional form of the covariance. Additionally, \celerite 
-	only allows for one-dimensional GPs.
+	to $N$. \celerite achieves this increased speed by limiting the user to GPs with a specific functional form of the covariance, albeit a useful and 
+	versatile one. Additionally, the \celerite method is only applicable to one-dimensional GPs. Here we present a method to compute a two-dimensional 
+	GP on a grid in $\mathcal{O}(NM^3J^2)$ time for an $N\times M$ grid when the covariance in one of the dimensions is described by the \celerite 
+	kernel. This is an improvement over the general case in which the computation runs in $\mathcal{O}(N^3M^3)$. Our method is especially useful 
+	in cases where $M << N$ and $M$ is constant. In this case the factor of $M^3$ in the runtime can be thought of as a constant. An example would 
+	be modeling variability in multiband photometry where $M$ would represent the number of bands and would be small compared to $N$.
 

### 2018-10-31 examples/occultquad.jl

@@ -1,4 +1,5 @@
-function occultquad(z0::Float64,u1::Float64,u2::Float64,p::Float64)
+
+function occultquad(z0,u1,u2,p)
 #
 # Julia implementation of Mandel & Agol (2002).
 # Computes the limb-darkened transit (or eclipse) of a star.
@@ -50,7 +51,7 @@ function occultquad(z0::Float64,u1::Float64,u2::Float64,p::Float64)
     etad = (kap1 + p^2*(p^2+2z^2)*kap0 - 0.25*(1+5p^2+z^2)*sqrt((1-a)*(b-1)))/(2pi)
     sqt=0.25*(4z^2-(1+z^2-p^2)^2)
     lambdae = (p^2*kap0+kap1-sqrt(sqt*(sqt >= 0)))/pi
-    return 1.0 - ((1.-u1-2u2)*lambdae + (u1+2u2)*(lambdad+2//3*(p > z)) + u2*etad)/omega
+    return 1.0 - ((1. - u1 - 2u2)*lambdae + (u1+2u2)*(lambdad+2//3*(p > z)) + u2*etad)/omega
 #    println(z,' ',"Case 2")
   end
 # Cases 3/9: planet within disk of star
@@ -64,7 +65,7 @@ function occultquad(z0::Float64,u1::Float64,u2::Float64,p::Float64)
     etad = 0.5*p^2*(p^2+2z^2)
 #    println("etad ",etad)
     lambdae = p^2
-    return 1.0 - ((1.-u1-2u2)*lambdae + (u1+2u2)*(lambdad+2//3*(p > z)) + u2*etad)/omega
+    return 1.0 - ((1. -u1 - 2u2)*lambdae + (u1+2u2)*(lambdad+2//3*(p > z)) + u2*etad)/omega
 #    println("lambdae ",lambdae)
 #    println(z,' ',"Cases 3/9")
   end
@@ -73,7 +74,7 @@ function occultquad(z0::Float64,u1::Float64,u2::Float64,p::Float64)
     lambdad = 2//3/pi*acos(1-2p)-4//9/pi*(3+2p-8p^2)*sqrt(p*(1-p))-2/3*(p > 0.5)
     etad = 0.5*p^2*(p^2+2z^2)
     lambdae = p^2
-    return 1.0 - ((1.-u1-2u2)*lambdae + (u1+2u2)*(lambdad+2//3*(p > z)) + u2*etad)/omega
[…]

### 2018-10-31 examples/transit.jl

@@ -1,0 +1,17 @@
+include("transit_orb.jl")
+
+function transit(t, x)
+    nsub = 5
+    n = length(t)
+    flux::Array{Number, 1} = zeros(n)
+    for i=1:n
+        ti = t[i]
+        if i > 1
+            dt = ti-t[i-1]
+        else
+            dt = 1e-10
+        end
+        flux[i] = transit_orb(ti, x, dt, nsub)
+    end
+    return flux
+end

### 2018-10-31 examples/transit_orb.jl

@@ -1,3 +1,5 @@
+include("kepler.jl")
+include("occultquad.jl")
 function transit_orb(t,x,dt,nsub)
 # Computes a transit lightcurve, normalized to unity, as a
 # function of time, t, usually in HJD or BJD.
@@ -18,35 +20,27 @@ function transit_orb(t,x,dt,nsub)
 f1=1.5*pi-x[10]*pi/180.0
 ecc=x[9]
 tp=(x[4]+x[1]*sqrt(1.0-ecc^2)/2.0/pi*(ecc*sin(f1)/(1.0+ecc*cos(f1))
-    -2.0/sqrt(1.0-ecc^2)*atan2(sqrt(1.0-ecc^2)*tan(0.5*f1),1.0+ecc)))
-flux_prior = -1.0
+    -2.0/sqrt(1.0-ecc^2)*atan(sqrt(1.0-ecc^2)*tan(0.5*f1),1.0+ecc)))
 fluxoft = 0.0
-nsub = 1
-while abs(fluxoft-flux_prior) > 1e-10 
-  flux_prior = fluxoft
-  fluxoft = 0.0
-  for j=1:nsub
-    tsub = t-dt/2.0+dt*(j-0.5)/nsub
-    m=2.0*pi/x[1]*(tsub-tp)
-    f=kepler(m,ecc)
-    radius=x[8]*(1.0-ecc^2)/(1.0+ecc*cos(f))
+for j=1:nsub
+  tsub = t-dt/2.0+dt*(j-0.5)/nsub
+  m=2.0*pi/x[1]*(tsub-tp)
+  f=kepler(m,ecc)
+  radius=x[8]*(1.0-ecc^2)/(1.0+ecc*cos(f))
 # Now compute sky separation:
-    z0=radius*sqrt(1.0-(sin(x[2]*pi/180.0)*sin(x[10]*pi/180.0+f))^2)
-    if (sin(x[10]*pi/180.0+f) < 0.0) && (z0 <= (1.0+x[3]))
-      if x[3] < 1.0
-        fluxoft += occultquad(z0,x[5],x[6],x[3])
-      else
-# We'll assume that the smaller object is not limb-darkened:
-        fluxoft += occultquad(z0,  0.,  0.,x[3])
-      end 
+  z0=radius*sqrt(1.0-(sin(x[2]*pi/180.0)*sin(x[10]*pi/180.0+f))^2)
[…]

### 2018-10-31 src/gp.jl

@@ -5,27 +5,27 @@ using LinearAlgebra
 mutable struct Celerite
     kernel::Term
     computed::Bool
-    D::Vector{Float64}
-    W::Array{Float64}
-    up::Array{Float64}
-    phi::Array{Float64}
-    x::Vector{Float64}
-    logdet::Float64
+    D::Vector{Real}
+    W::Array{Real}
+    up::Array{Real}
+    phi::Array{Real}
+    x::Vector{Real}
+    logdet::Real
     n::Int
     J::Int
     
-    Q::Array{Float64, 2}
+    Q::Array{Real, 2}
 
 #    Celerite(kernel) = new(kernel, false, [], [], [], [], [])
 #    Celerite(kernel) = new(kernel, false, zeros(Float64,0),zeros(Float64,0),zeros(Float64,0,0), zeros(Float64,0,0), zeros(Float64,0,0), zeros(Float64,0,0), zeros(Float64,0),0.0,0,0)
     Celerite(kernel, Q=ones(1, 1)) = new(kernel, false, zeros(Float64,0),zeros(Float64,0,0),zeros(Float64,0,0),zeros(Float64,0,0),zeros(Float64,0),0.0,0,0,Q)
 end
 
-function cholesky_ldlt!(a_real::Vector{Float64}, c_real::Vector{Float64},
-                       a_comp::Vector{Float64}, b_comp::Vector{Float64}, 
-                       c_comp::Vector{Float64}, d_comp::Vector{Float64},
-                       t::Vector{Float64}, var::Vector{Float64}, X::Array{Float64,2}, 
-                       phi::Array{Float64,2}, u::Array{Float64,2}, D::Vector{Float64})
+function cholesky_ldlt!(a_real::Vector{<:Real}, c_real::Vector{<:Real},
+                       a_comp::Vector{<:Real}, b_comp::Vector{<:Real}, 
+                       c_comp::Vector{<:Real}, d_comp::Vector{<:Real},
+                       t::Vector{<:Real}, var::Vector{<:Real}, X::Array{<:Real,2}, 
+                       phi::Array{<:Real,2}, u::Array{<:Real,2}, D::Vector{<:Real})
 #
 # Fast LDLT Cholesky solver based on low-rank decomposition due to Sivaram, plus
 # real implementation of celerite term.
[…]

### 2018-10-31 src/terms.jl

@@ -280,9 +280,9 @@ end
 
 # A SHO term
 mutable struct SHOTerm <: Term
-    log_S0::Float64
-    log_Q::Float64
-    log_omega0::Float64
+    log_S0
+    log_Q
+    log_omega0
 end
 
 function get_real_coefficients(term::SHOTerm)

### 2018-12-19 paper/paper.tex
updated paper

--- 2018-10-19 paper/paper.tex
@@
 %% using aastex version 6.2
-\documentclass[modern]{aastex62}
+\documentclass[preprint2]{aastex62}
 
@@
 \begin{abstract}
-Gaussian processes are frequently used to model variability in astrophysical time-series. In addition to studying variability in stars and AGN, Gaussian 
-process regression can be used to account for correlated noise in light curves of exoplanetary transits and micro-lensing events. One of the chief limitations of this 
-technique is the computational time required to compute a Gaussian process model for large data-sets. This also limits prospects for using Gaussian processes with 
-multi-dimensional datasets, although sparse-matrix approximations can make GP regression on multi-dimensional datasets feasible for some applications. Here we 
-present a method based on the \celerite GP implementation which enables fast computation of two-dimensional gaussian process noise models for datasets with a 
-small second dimension, such as multi-band photometry. This method enables the use of GP regression with multi-wavelength transit and micro-lensing light curves, 
-and may have applications to radial velocity and direct imaging data as well.
-
+Gaussian processes are a model of stochastic variability frequently used in analyzing astrophysical time-series. 
+They have been used to study variability in light curves of stars and AGN, as well as the spatial 
+distribution of the extragalactic dust distribution. In the realm of stellar variability, GP regression 
+is used to characterise stellar rotation periods, search for transit signals in noisy data, and analyze 
+RV curves. One of the chief limitations of this method is the computational time incurred in computing 
+a GP model for large or multi-dimensional data-sets. While approximate methods can make GP regression 
+feasible for some of these applications, in others exact methods are preferable. Here 
+we present a method based on the \celerite GP implementation which enables fast, exact 
+computation of two-dimensional GP models for data-sets with a small second dimension. This includes 
+multi-bandpass photometry of transit and microlensing light curves. Our method also may have applications 
+to RV and direct imaging methods. 
 \end{abstract}
@@
 %% to the KEY in the \bibitem in the reference list below. 
-\tableofcontents
+%\tableofcontents
+\section{introduction}\label{sec:intro}
+    A Gaussian process (GP) is a generalization of a Gaussian distribution to 
+    function-space \citep{Rasmussen2006}. GPs have been used extensively in 
+    astrophysics as a model of stochastic variability across both space and time, 
+    to study a diverse range of phenomena including variable stars, AGN\citep{Kelly2014}, the 
+    extragalactic dust distribution\citep{Sale2018}, and gravitational wave signals\citep{Moore2016}. 
+    They have also been used to account 
+    for correlated noise in photometric and RV measurements of 
+    exoplanet hosts\citep{Aigrain2016,Jones2017,Rajpaul2015}.
 
-\section{introduction}\label{sec:intro}
-	Gaussian processes have been used extensively in astrophysics to model stochastic variability across space and time. 
-	They have been used to study variable stars and AGN, to model the extragalactic distribution of dust, and to model correlated noise in 
-	photometric and radial velocity observations of exoplanet host stars.	\gordontodo{X-ray binaries?}
+    GP models offer advantages over other methods of modeling stochastic variability 
+    because they are able to account for correlated variations without the imposition 
+    of a parametric model. \gordontodo{more about advantages of GPs}
 
-	\gordontodo{The following paragraph might make more sense in the exomoons section.}
-	Gaussian process noise models are most advantageous when correlated noise dominates over white noise. 
-	This is already the case for state-of-the-art radial velocity observations, and transit 
-	photometry is not far behind. Future space-based observatories such as James Webb 
-	will enable high-precision photometry in which stellar variability is the main barrier to detecting ever more shallow transit signals. 
-	Stellar variability which obscures transit signal can be a result of spot and faculae crossings, asteroseismic oscillations, or variable 
-	granulation on the star's surface. GPs are able to \gordontodo{something about how a GP model is non-parametric and the advantages thereof? Also agnostic 
-	to the source of variability (seems related)}
+    The primary limitation of GP methods is that most GP operations are 
+    computationally expensive. For instance, the common tasks 
+    of computing likelihoods and sampling from the GP require 
+    $\mathcal{O}(N^3)$ operations for $N$ data points. This is 
+    especially problematic for use cases which require repeated calls 
+    to the likelihood function, such as minimization and performing 
+    MCMC. This has prevented the adoption of GP methods for data-sets 
+    larger than $N\sim 10^4$ points\citep{Deisenroth2015}.
 
-	\gordontodo{This is where we come in...}
+    The issue of computational expense can be partially overcome either by approximating 
+    the GP or by restricting the user to a subset GPs for which the computation 
+    can be sped up\citep{Rasmussen2006}. In the latter category we 
+    find \celerite\citep{Foreman-Mackey2017}, a fast one-dimensional GP 
+    method which increases the speed of most common GP computations to $\mathcal{O}(NJ^2)$ 
+    where $J$ is typically small compared to $N$. \celerite achieves this speedup by 
+    limiting the user to GPs with a specific functional form of the covariance. 
+    Another limitation of \celerite is that the method is only applicable to 
+    one-dimensional GPs. Here we present a method to rectify this difficiency of 
+    the original \celerite method. Our extension of \celerite computes two-dimenional 
+    GPs on an $N$ by $M$ grid in $\mathcal{O}(NJ^2M^3)$ operations. This is an improvement 
+    over the general case in which the computation of the same GP would require $\mathcal{O}(N^3M^3)$ 
+    operations. Because the speed of our computation is still sharply dependent on $M$, it 
+    is most useful in cases where $M << N$. An example would be modeling variability in 
+    multi-bandpass photometry where $M$ would represent the number of bands.
 
-	\gordontodo{Are GPs applicable to instrumental noise as well? Seems best to account for instrumental effects with models that account for known factors since 
-	they might not be well-described by a gaussian process, right?}
-
-	\gordontodo{What about other methods of accounting for astrophysical noise? i.e. cosine filtering, polynomials, etc. --- also related to GP as non-parametric 
-	noise model}
+    In this paper we first review the \celerite method before introducing our extension 
+    of the original method. We then give a detailed example of GP regression on multi-band 
+    transit photometry to recover transit parameters in the presence of significant 
+    correlated noise. Finally, we discuss other potential applications to our method and 
+    conclude with an outline of future work in which we hope to further reduce the 
+    computational expense of a two-dimensional GP. This would enable the use of 
+    GP methods on even larger data-sets and potentially make feasible GP computations 
+    in higher dimensions.
 
@@
 	\begin{equation} 
-		\bvec{\mu_\theta}(\bvec{x}) = \expandvec{\mu_\bvec{\theta}(x_1)}{\mu_\bvec{\theta}(x_N)}
+		\bvec{\mu_\theta}(\bvec{x}) = \expandvec{\mu_\bvec{\theta}(x_1)}
+		    {\mu_\bvec{\theta}(x_N)}
 	\end{equation}
-	and covariance defined the kernel function $k_\bvec{\alpha}(x_n, x_m)$. The mean and kernel functions depend on the parameters 
-	$\bvec{\theta}$ and $\bvec{\alpha}$ respectively, which are referred to as ``hyperparameters" of the GP. The kernel function defines a covariance matrix 
-	$\left[K\right]_{n, m} = k_\alpha(x_n, x_m)$. This model has a likelihood function given by
-	\begin{equation}
-		\ln\ \mathcal{L(\bvec{\theta}, \bvec{\alpha})} = \ln\ p(\bvec{y}|X, \bvec{\theta}, \bvec{\alpha}) = 
-			-\frac{1}{2}\bvec{r_\theta}^\T K_\bvec{\alpha}^{-1}\bvec{r_\theta} 
-			-\frac{1}{2}\ln\ \mathrm{det}(K_\bvec{\alpha}) - \frac{N}{2}\ln(2\pi)
-	\end{equation}
-	\gordontodo{I'm using $X$ here because DFM does, but wouldn't $\bvec{x}$ be more clear?}
+	and covariance defined the kernel function $k_\bvec{\alpha}(x_n, x_m)$. 
+	The mean and kernel functions depend on the parameters 
+	$\bvec{\theta}$ and $\bvec{\alpha}$ respectively, which are referred to 
+	as ``hyperparameters" of the GP. The kernel function defines a covariance matrix 
+	$\left[K\right]_{n, m} = k_\alpha(x_n, x_m)$. This model has a likelihood 
+	function given by
+	\begin{eqnarray}
+		\ln\ \mathcal{L(\bvec{\theta}, \bvec{\alpha})} &=& \ln\ 
+		    p(\bvec{y}|\bvec{x}, \bvec{\theta}, \bvec{\alpha}) \\ 
+		&=& -\frac{1}{2}\bvec{r_\theta}^\T K_\bvec{\alpha}^{-1}\bvec{r_\theta} 
+			-\frac{1}{2}\ln\ \mathrm{det}(K_\bvec{\alpha}) - \frac{N}{2}\ln(2\pi) \nonumber
+	\end{eqnarray}
 	where 
@@
 	\end{equation}
-	By maximizing the likelihood (or, in practice, minimizing the negative of the log-likelihood), one can obtain an estimate of the hyperparameters $\bvec{\theta}$ and 
-	$\bvec{\alpha}$. Furthermore, Markov Chain Monte Carlo (MCMC) methods can be used to sample the likelihood in hyperparameter-space. \gordontodo{expand 
-	on this.}
+	By maximizing the likelihood (or, in practice, minimizing the negative 
+	of the log-likelihood), one can obtain an estimate of the hyperparameters 
+	$\bvec{\theta}$ and $\bvec{\alpha}$. Markov Chain Monte Carlo (MCMC) methods can be used to 
+	sample the posterior distribution over the hyperparameters. 
 	\subsection{The \celerite model}
-		The primary limitation of Gaussian process noise models is that they are computationally expensive, with computation of the likelihood function scaling as 
-		$\mathcal{O}(N^3)$ due to the matrix inversion. This is especially problematic for use cases such as MCMC which require calling the likelihood function many 
-		times. This has prevented the adoption of Gaussian process noise models for datasets larger than $N \approx 10^3$, except in cases where approximate 
-		methods suffice. The \celerite method overcomes this issue by reducing the time to compute the likelihood function to $\mathcal{O}(NJ^2)$ for one 
-		class of kernel functions; where J is the number of \celerite terms making up the kernel function of the GP. This class of kernel functions are given by 
+		The primary limitation of Gaussian process noise models is that they 
+		are computationally expensive, with computation of the likelihood function 
+		scaling as $\mathcal{O}(N^3)$ due to the matrix inversion. This is especially 
+		problematic for use cases such as MCMC which require calling the likelihood 
+		function many times. This has prevented the adoption of Gaussian process noise models for 
+		datasets larger than $N \approx 10^3$, except in cases where approximate 
+		methods suffice. The \celerite method overcomes this limitation by reducing 
+		the time to execute most GP computations to linear time for one class of 
+		kernel functions\citep{Foreman-Mackey2017}. We give a brief description of the \celerite method 
+		and its limitations below.
+
+		Consider a one-dimensional gaussian process evaluated at the coordinates
+		\begin{equation}
+			\bvec{x} = \expandvec{t_1}{t_N}
+		\end{equation}
+		The \celerite kernel is given by
 		\begin{equation}
@@
 		\end{equation}
-		where $\bvec{\alpha} = (a_1...a_j, c_1...c_j)$, $\sigma_n^2$ is the variance of the white noise, and $\tau_{nm} = |t_n-t_m|$. 
-		The coefficients $a$ and $c$ may be complex.
+		where $\bvec{\alpha} = (a_1...a_j, c_1...c_j)$, $\sigma_n^2$ is the variance of the gaussian-distributed white noise, and $\tau_{nm} = |t_n-t_m|$. 
+		The coefficients $a$ and $c$ may be complex, in which case we introduce $b$ and $d$ which are the imaginary parts of $a$ and $c$ respectively. 
+		This kernel defines a \celerite model with $J$ terms.
 
-		For kernel functions of this form, the covariance matrix can be written as a semi-separable matrix of rank $2J$:
+		For a kernel function of this form, the covariance matrix is a symmetric, semiseparable with semiseparability rank $R=2J$. A matrix of this type can 
+		be written in terms of two generator matrices $U$ and $V$, both of size $(N\times 2J)$, and a diagonal matrix $A$:
 		\begin{equation}
+		\label{eqn:KUV}
 			K = A + \mathrm{tril}(UV^\T) + \mathrm{triu}(VU^\T)
 		\end{equation}	
-		where $U$ and $V$ have entries given by: 
+		In the case of our covariance matrix, the generator matrices are specified by:	
 		\begin{eqnarray}
@@
 		\end{eqnarray}
-		and A is a diagonal matrix:
+		and A is given by:
 		\begin{equation}
@@
 		\end{equation}
-		Cholesky decomposition can be performed on this matrix in $\mathcal{O}(NJ^2)$ time, as described in the appendix.
+		Computing the Cholesky decomposition for this covariance matrix can be accomplished in $\mathcal{O}(NJ^2)$ operations, allowing for the 
+		fast computation of the GP likelihood function\footnote{The algorithm is described in appendex \ref{sec:celerite_algorithm}}.
 
-		\gordontodo{Add discussion of the versatility of this kernel, and it's connection to stochastically driven harmonic oscillators.}
+		%\gordontodo{Add discussion of the versatility of this kernel, and it's connection to stochastically driven harmonic oscillators.}
+		%
+		%\gordontodo{Add discussion of semiseperability --- the reader should understand that the limitation of \celerite is that only kernels of this form can be 
+		%represented by a semi-seperable matrix, which is what makes moving to two dimensions slightly more complicated than ordinary.}
+		The main limitation of \celerite is that the method only applies to one class of kernels. However, the \celerite kernel overcomes this limitation 
+		by being both versatile in its ability to approximate other kernels, and also directly applicable to many problems. The versatility of this 
+		kernel is demonstrated in \gordontodo{figure reference} which shows approximations to several popular kernels achieved by appropriate choices 
+		of the \celerite coefficients $a_j$ and $c_j$.
 
-		\gordontodo{Add discussion of semiseperability --- the reader should understand that the limitation of \celerite is that only kernels of this form can be 
-		represented by a semi-seperable matrix, which is what makes moving to two dimensions slightly more complicated than ordinary.}
+		One model of stellar oscillations models the star as a stochastically-driven damped harmonic oscillator\citep{Anderson1990}. The solution to the equations describing 
+		such an oscillator is a GP with a \celerite kernel with coefficients 
+		\begin{eqnarray}
+			a_{j\pm} &=& \frac{1}{2}S_0\omega_0\left[1\pm\frac{1}{\sqrt{1-4Q^2}}\right] \\
+			c _{j\pm}&=& \frac{\omega_0}{2Q}\left[1 \mp \sqrt{1-4Q^2}\right]
+		\end{eqnarray}
+		where $S_0$ is proportional to the power at the un-damped frequency of the oscillator, $\omega_0$, and $Q$ is a parameter known as 
+		the quality factor of the oscillator. Variations in the stellar radius due to these oscillations gives rise to brightness variations in the star's 
+		light curve \gordontodo{reference?}. 
+		This means that a \celerite GP can be used as a model for stellar variability due 
+		to p-mode oscillations. Additionally, setting $Q=1/\sqrt{2}$ results in a PSD 
+		which has been used to describe granulation-driven stellar variability\citep{Kallinger2014}.
 
 \section{2d gaussian processes}
-	\subsection{kronecker structure of the 2d covariance matrix}
-	In two dimensions the kernel function depends on two coordinates. We will continue to use
+	In addition to being limited to one class of kernels \celerite, as formulated above, is inherently one-dimensional. For many applications, such 
+	as modeling noise in single-bandpass light curves or RV time series GP computations in one dimension are all that is necessary. However in 
+	other cases such as modeling noise in simultaneous multi-bandpass light curves or disentangling noise in RV data using multiple activity 
+	indicators, two-dimensional GP methods are necessary. We outline a method to compute a GP in two dimensions utilizing the \celerite 
+	method for the larger of the two dimensions to speedup most GP operations by a factor of $\sim\mathcal{O}(N^2)$ over ordinary 2D GPs.
+
+	In two-dimensions the independent variable is
 	\begin{equation}
-		\bvec{t} = \expandvec{t_1}{t_N}
+		\bvec{x} = \expandvec{(t_1, u_1)}{(t_N, u_N)}
 	\end{equation}
-	to refer to the first coordinate, and introduce
+	If the covariance between two points satisfies 
 	\begin{equation}
-		\bvec{u} = \expandvec{u_1}{u_M}
+	\label{eqn:multiplication_condition}
+		k_\alpha(x_n, x_m) = q_\alpha(t_n, t_m)r_\alpha(u_n, u_m)
 	\end{equation}
-	to refer to the second coordinate. In most cases, we will demand that $M << N$. We then define $\bvec{x}$ to be the list of points
+	then the covariance matrix has the structure of a Hadamard matrix product (element-wise 
+	multiplication )
 	\begin{equation}
-		\bvec{x} = \expandvec{(t_1, u_1), (t_1, u_2), \cdots (t_1, u_M)}{(t_N, u_M)}
+		\left[K_\alpha\right]_{nm} = \left[Q_\alpha\right]_{nm}\left[R_\alpha\right]_{nm}
 	\end{equation}
-	We can also write this as
+	or 
 	\begin{equation}
-		x_i = (t_{\lfloor(i-1)/M\rfloor + 1}, u_{i - \lfloor(i-1)/M\rfloor M})
+		K = Q\circ R
 	\end{equation}
-	The covariance between $x_i$ and $x_j$ is then given by
+	Now consider a GP evaluated on a $(N\times M)$ grid. We begin by defining
+	\begin{eqnarray}
+		\bvec{t'} &=& \expandvec{t'_1}{t'_N} \\
+		\bvec{u'} &=& \expandvec{u'_1}{u'_M}
+	\end{eqnarray}
+	to be vectors containing only the unique $t$ and $u$ values in increasing order. 
+	%, such that 
+	%\begin{equation}
+	%	\bvec{x} = \expandvec{((t'_1, u'_1)\cdots(t'_N,u'_1), (t'_1, u'_2)\cdots(t'_N,u'_2)}{(t'_1, u'_M)\cdots(t'_N, u'_M)}
+	%\end{equation}
+	We also define the corresponding covariance matrices for each dimension: 
+	\begin{subequations}
+	\label{eqn:QR_defs}
+	\begin{eqnarray}
+		\left[Q'_\alpha\right]_{nm} &=& q_\alpha(t'_n, t'_m) \\
+		\left[R'_\alpha\right]_{pq} &=& r_\alpha(u'_p, u'_q) 
+	\end{eqnarray}
+	\end{subequations}
+	In this case, assuming (\ref{eqn:multiplication_condition}), the full covariance matrix can be written as a Kronecker product between these 
+	two one-dimensional covariances\footnote{See appendix \ref{sec:matrix_prod} for a proof of (\ref{eqn:kronecker_K}) and notes on the Hadamard and Kronecker products. }: 
 	\begin{equation}
-		k_{\bvec{\alpha}\bvec{\beta}}(x_i, x_j) = \sigma_i^2\delta_{ij} + q_\bvec{\alpha}(\tau_{ij})\circ r_\bvec{\beta}(\nu_{ij})
+	\label{eqn:kronecker_K}
+		K_\alpha = Q'_\alpha \otimes R'_\alpha
 	\end{equation}
-	where $\tau_{ij} = |t_{\lfloor(i-1)/M\rfloor + 1}-t_{\lfloor(j-1)/M\rfloor + 1}|$, 
-	$\nu_{ij} = |u_{i - \lfloor(i-1)/M\rfloor M}-u_{j - \lfloor(j-1)/M\rfloor M}|$; 
-	$q$ and $r$ are one-dimensional kernel functions which depend on the hyperparameters $\bvec{\alpha}$ and 
-	$\bvec{\beta}$ respectively; 
-	and $\circ$ is some binary operation. 
-	If $Q_\bvec{\alpha}$ and $R_\bvec{\alpha}$ are the covariance matrices corresponding to 
-	$q_\bvec{\alpha}$ and $r_\bvec{\alpha}$ such that
+	Since the rest of this discussion will focus on the case of gridded data, we will drop the prime notation and use $Q_\alpha$ and $R_\alpha$ to refer to the 
+	matrices in (\ref{eqn:QR_defs}). In the following discussion we assume that the covariance in $t$ is described by a \celerite kernel, and that 
+	the covariance in $u$ is arbitrary. This means that $Q_\alpha$ can be expressed in terms of the generator matrices $U$ and $V$.  Following equation 
+	(\ref{eqn:KUV}), in the absence of a white noise term, we can write the covariance matrix as follows\footnote{See appendix \ref{sec:Kproof} for proof}:
+	\begin{eqnarray}
+	\label{eqn:KUVR}
+		%\nonumber K &=& A\otimes R + \mathrm{tril}(UV^\T)\otimes R + \mathrm{triu}(VU^\T)\otimes R \\
+		%\nonumber &=& A\otimes R + \mathrm{tril}((UV^\T)\otimes (RI_M)) + \mathrm{triu}((VU^\T)\otimes (RI_M) ) \eqnum{\value{equation}} \\
+		\nonumber K &=& \left(A + \mathrm{tril}(UV^\T) + \mathrm{triu}(VU^\T)\right)\otimes R \\
+		&=& A' + \mathrm{tril}(U'V'^\T) + \mathrm{triu}(V'U'^T)
+	\end{eqnarray}
+	where
+	\begin{eqnarray}
+		A' &=& A\otimes R \\
+		U' &=& U\otimes R \\
+		V' &=& V\otimes I_M
+	\end{eqnarray}
+	To include white noise we add a white noise term to each element of A:
 	\begin{equation}
-		\left[Q_\bvec{\alpha}\right]_{ij} = q_\bvec{\alpha}(\tau_{ij})\ \ \ \mathrm{and}\ \ \ \left[R_\bvec{\beta}\right]_{ij} = r_\bvec{\beta}(\tau_{ij})
+		A'_{n, n} \to A'_{n, n} + \sigma_n^2
 	\end{equation}
-	then, using $q_{ij}$ and $r_{ij}$ as shorthand for the entries in $Q_\bvec{\alpha}$ and $R_\bvec{\beta}$, $K_\bvec{\alpha\beta}$ can be represented as: 
-	\begin{equation}
-		K_\bvec{\alpha\beta} = \Sigma + 
-			\begin{bmatrix}
-				q_{11}\circ R_\bvec{\beta} & ... & q_{1N}\circ R_\bvec{\beta} \\
-				\vdots & \ddots & \\
-				q_{N1}\circ R_\bvec{\beta} & ... & q_{NN}\circ R_\bvec{\beta}
-			\end{bmatrix}
-	\end{equation}
-	where $[\Sigma]_{ij} = \sigma_i^2\delta_{ij}$ and
-	\begin{equation}
-		q_{pq}\circ R_\bvec{\alpha} = 
-			\begin{bmatrix}
-				q_{pq}\circ r_{11} & ... & q_{pq}\circ r_{1M} \\
-				\vdots & \ddots & \\
-				q_{pq}\circ r_{M1} & ... & q_{pq}\circ r_{MM}
-			\end{bmatrix}
-	\end{equation}
-	If the operator $\circ$ represents multiplication, we have 
-	\begin{equation}
-		K_\bvec{\alpha\beta} = \Sigma + Q_\bvec{\alpha} \otimes  R_\bvec{\beta}
-	\end{equation}
-	where $\otimes$ is the Kronecker product. If the covariance in $\bvec{t}$ is described by a \celerite kernel can write 
-	\begin{eqnarray}
-		K &=& \Sigma + \left[\mathrm{tril}(UV^\T) + \mathrm{triu}(VU^\T)\right]\otimes R \\
-		&=& \Sigma + \mathrm{tril}\left[ (U\otimes R)(V\otimes I_M)^\T\right] + \mathrm{triu}\left[ (V\otimes I_M)(U\otimes R)^\T\right]
-	\end{eqnarray}
-	This covariance has the same semi-separable form as in the one-dimensional case, but now $K$ is an $(NM\times NM)$ matrix of rank $2JM$. 
-	This means that we can apply the same Cholesky decomposition algorithm (described in the appendix) as in the 1d case to compute the GP in 
-	$\mathcal{O}(NJ^2M^3)$. While our implementation only handles the Kronecker product case, 
-	his result holds in the case that $\circ$ is an operation other than multiplication, so long as $\circ$ has the property
-	\begin{equation}
-		(a\circ B)(c\circ D) = (ac)\circ(BD)
-	\end{equation}
-	for scalars $a$ and $b$ and matrices $B$ and $D$. 
+	From the form of equation \ref{eqn:KUVR} we can see that the covariance is, as before, semiseparable. However Kronecker 
+	multiplication by $R$ and $I_M$ has increased the size of the generator matrices from $(N\times 2J)$ to $(NM\times 2JM)$. As a result,  
+	the number of operations necessary to compute the Cholesky decomposition to increase from $\mathcal{O}(NJ^2)$ to $\mathcal{O}(NJ^2M^3)$. 
+	The number of operations for sampling from and extrapolating/interpolating with the two-dimensional GP scales the same. Some improvement 
+	can be made in the scaling for extrapolation and interpolation if the covariance in the second dimension can be expressed via 
+	a \celerite kernel. In this case the algorithm requires $\mathcal{O}(nNP + mMP)$ operations where $P$ is the number of points at which the prediction is evaluated and $n, m$ are constants. The 
+	Cholesky composition algorithm as well as algorithms for interpolation/extrapolation and sampling from the GP are described in appendix 
+	\ref{sec:algorithms}.
+
+	It should be noted that the scaling of our method is still relatively poor as we add points to our grid in the second dimension. This 
+	is why the method is most suitable for cases where the second dimension is much smaller than the first, i.e. where $M << N$. 
+	This condition applies to the important cases of modeling noise in multi-bandpass observations. 
+	%Into appendix? 
+	%So that $Q'_\alpha$ is a symmetric $(N\times N)$ matrix (with hyperparameters $\alpha$) 
+	%describing the covariance for a one-dimensional GP evaluated along $\bvec{t}$ and 
+	%$R'_\alpha$ is the same for $\bvec{u}$.  These matrices can be related back to the un-primed $Q_\alpha$ and $R_\alpha$ 
+	%matrices via the Kronecker product.
+
+	%\subsection{kronecker structure of the 2d covariance matrix}
+	%In two dimensions the kernel function depends on two coordinates. We will continue to use
+	%\begin{equation}
+	%	\bvec{t} = \expandvec{t_1}{t_N}
+	%\end{equation}
+	%to refer to the first coordinate, and introduce
+	%\begin{equation}
+	%	\bvec{u} = \expandvec{u_1}{u_M}
+	%\end{equation}
+	%to refer to the second coordinate. In most cases, we will demand that $M << N$. We then define $\bvec{x}$ to be the list of points
+	%\begin{equation}
+	%	\bvec{x} = \expandvec{(t_1, u_1), (t_1, u_2), \cdots (t_1, u_M)}{(t_N, u_M)}
+	%\end{equation}
+	%We can also write this as
+	%\begin{equation}
+	%	x_i = (t_{\lfloor(i-1)/M\rfloor + 1}, u_{i - \lfloor(i-1)/M\rfloor M})
+	%\end{equation} 
+	%The covariance between $x_i$ and $x_j$ is then given by
+	%\begin{equation}
+	%	k_{\bvec{\alpha}\bvec{\beta}}(x_i, x_j) = \sigma_i^2\delta_{ij} + q_\bvec{\alpha}(\tau_{ij})\circ r_\bvec{\beta}(\nu_{ij})
+	%\end{equation}
+	%where $\tau_{ij} = |t_{\lfloor(i-1)/M\rfloor + 1}-t_{\lfloor(j-1)/M\rfloor + 1}|$, 
+	%$\nu_{ij} = |u_{i - \lfloor(i-1)/M\rfloor M}-u_{j - \lfloor(j-1)/M\rfloor M}|$; 
+	%$q$ and $r$ are one-dimensional kernel functions which depend on the hyperparameters $\bvec{\alpha}$ and 
+	%$\bvec{\beta}$ respectively; 
+	%and $\circ$ is some binary operation. 
+	%If $Q_\bvec{\alpha}$ and $R_\bvec{\alpha}$ are the covariance matrices corresponding to 
+	%$q_\bvec{\alpha}$ and $r_\bvec{\alpha}$ such that
+	%\begin{equation}
+	%	\left[Q_\bvec{\alpha}\right]_{ij} = q_\bvec{\alpha}(\tau_{ij})\ \ \ \mathrm{and}\ \ \ \left[R_\bvec{\beta}\right]_{ij} = r_\bvec{\beta}(\tau_{ij})
+	%\end{equation} 
+	%then, using $q_{ij}$ and $r_{ij}$ as shorthand for the entries in $Q_\bvec{\alpha}$ and $R_\bvec{\beta}$, $K_\bvec{\alpha\beta}$ can be represented as: 
+	%\begin{equation}
+	%	K_\bvec{\alpha\beta} = \Sigma + 
+	%		\begin{bmatrix}
+	%			q_{11}\circ R_\bvec{\beta} & ... & q_{1N}\circ R_\bvec{\beta} \\
+	%			\vdots & \ddots & \\
+	%			q_{N1}\circ R_\bvec{\beta} & ... & q_{NN}\circ R_\bvec{\beta}
+	%		\end{bmatrix}
+	%\end{equation}
+	%where $[\Sigma]_{ij} = \sigma_i^2\delta_{ij}$ and
+	%\begin{equation}
+	%	q_{pq}\circ R_\bvec{\alpha} = 
+	%		\begin{bmatrix}
+	%			q_{pq}\circ r_{11} & ... & q_{pq}\circ r_{1M} \\
+	%			\vdots & \ddots & \\
+	%			q_{pq}\circ r_{M1} & ... & q_{pq}\circ r_{MM}
+	%		\end{bmatrix}
+	%\end{equation}
+	%If the operator $\circ$ represents multiplication, we have 
+	%\begin{equation}
+	%	K_\bvec{\alpha\beta} = \Sigma + Q_\bvec{\alpha} \otimes  R_\bvec{\beta}
+	%\end{equation}
+	%where $\otimes$ is the Kronecker product. If the covariance in $\bvec{t}$ is described by a \celerite kernel can write 
+	%\begin{eqnarray}
+	%	K &=& \Sigma + \left[\mathrm{tril}(UV^\T) + \mathrm{triu}(VU^\T)\right]\otimes R \\
+	%	&=& \Sigma + \mathrm{tril}\left[ (U\otimes R)(V\otimes I_M)^\T\right] + \mathrm{triu}\left[ (V\otimes I_M)(U\otimes R)^\T\right]
+	%\end{eqnarray}
+	%This covariance has the same semi-separable form as in the one-dimensional case, but now $K$ is an $(NM\times NM)$ matrix of rank $2JM$. 
+	%This means that we can apply the same Cholesky decomposition algorithm (described in the appendix) as in the 1d case to compute the GP in 
+	%$\mathcal{O}(NJ^2M^3)$. While our implementation only handles the Kronecker product case, 
+	%his result holds in the case that $\circ$ is an operation other than multiplication, so long as $\circ$ has the property
+	%\begin{equation}
+	%	(a\circ B)(c\circ D) = (ac)\circ(BD)
+	%\end{equation}
+	%for scalars $a$ and $b$ and matrices $B$ and $D$. 
 	%where $a\circ B$ is the matrix 
@@
 	%\end{equation}
-	\gordontodo{See appendix for proof}
+	%\gordontodo{See appendix for proof}
 
-	\subsection{sampling the distribution}
-	A gaussian process can be used to simulate correlated noise. If $L$ is the lower triangular factor from the Cholesky decomposition of the 
-	$(N\times N)$ covariance matrix $K$, and $\bvec{z}$ is a vector of length $N$ containing random numbers drawn for a standard normal distribution, i.e.
+	%\subsection{sampling the distribution}
+	%A gaussian process can be used to simulate correlated noise. If $L$ is the lower triangular factor from the Cholesky decomposition of the 
+	%$(N\times N)$ covariance matrix $K$, and $\bvec{z}$ is a vector of length $N$ containing random numbers drawn for a standard normal distribution, i.e.
+	%\begin{equation}
+	%	\bvec{z} \sim \mathcal{N}(1, 0)
+	%\end{equation}
+	%then $\bvec{x} = \bvec{\mu_\bvec{\theta}(\bvec{x})} + L\bvec{z}$ is a vector with entries correlated by $K$ and with mean $\bvec{\mu_\theta}$, i.e.
+	%\begin{equation}
+	%	\bvec{x} \sim \mathcal{N}(\bvec{\mu_\theta}, K)
+	%\end{equation}
+	%\gordontodo{fix this so that $\mu$ depends on $\theta$ and is a function of $\bvec{t}$ or something like that. Compuational cost should be $\mathcal{O}(NJM^2)$}
+	%
+	%\subsection{interpolation and extrapolation}	
+	%\gordontodo{Bit more complicated --- if both dimensions are celerite processes then the scaling is better than in the case where $R$ is not celerite. You have notes on this.}
+
+%\section{example: exomoons}
+%	\gordontodo{What if I used Kepler-1625 as an example? Could simulate WFC3 GRISM data and show that moon could be validated with Celerite2D?}
+%	Gaussian process noise models are most advantageous when correlated noise dominates over white noise. 
+%	This is already the case for state-of-the-art radial velocity observations, and transit 
+%	photometry is not far behind. Future space-based observatories such as James Webb 
+%	will enable high-precision photometry in which stellar variability is the main barrier to detecting ever more shallow transit signals. 
+%	Stellar variability which obscures transit signal can be a result of spot and faculae crossings, asteroseismic oscillations, or variable 
+%	granulation on the star's surface.
+
+%	\gordontodo{Are GPs applicable to instrumental noise as well? Seems best to account for instrumental effects with models that account for known factors since 
+%	they might not be well-described by a gaussian process, right?}
+
+%	\gordontodo{What about other methods of accounting for astrophysical noise? i.e. cosine filtering, polynomials, etc. --- also related to GP as non-parametric 
+%	noise model}
+\section{example: solar system transits}
+    One application of fast two-dimensional Gaussian Processes is modeling noise in multi-bandpass light 
+    curves to, for example, search for transits and infer their parameters. One-dimensional GPs are already in use as 
+    noise models for transit light curves. By setting the mean of the GP equal to a transit model and minimizing the GP likelihood, 
+    transit parameters can be more accurately inferred in the presence of correlated stellar noise. However, for cases where a shallow 
+    transit signal is obscured by correlated noise, a one-dimensional GP noise model is not sufficient to disentangle signal from noise.
+
+    This problem can be mitigated by examining the wavelength-dependence of the star's variability. The correlated noise component 
+    of the light curve will be subject to this wavelength-dependence whereas the mean (the transit signal) will not share the same 
+    dependence. Thus by including multiple bandpasses and simultaneously modeling the covariance between bands using a two-dimensional 
+    GP.
+
+    In what follows we aim to demonstrate this method using transits of Solar System planets as an example. We use publicly available 
+    light curves from the SOHO mission's three-channel sunphotometer (SPH) \citep{Frohlich1995, Frohlich1997, Jimenez2002} and add Gaussian noise to simulate observations 
+    taken at varying distances from the Sun. We've injected transits of the Earth and Venus into that data and we demonstrate 
+    the recovery of transit parameters for both objects. 
+    \subsection{modeling the wavelength-dependence of stellar variability}
+        Our first task is to construct a model of the Sun's variability that allows us to determine the form of the $R$ matrix.
+The sun's p-mode oscillations peak at between 5 and 6 minutes \citep{Kiefer2018} compared to transit durations for Venus and Earth
+        of 11 and 13 hours respectively, so we ignore this source of variability in our model. Of more relevance are brightness variations modulated by changing patterns of 
+        hotter and cooler regions of the photosphere. We model the solar photosphere as 
+        a mix of two regions of temperatures $T_h$ and $T_c$ with $T_h > T_c$. The flux 
+        in a filter $A$ is 
         \begin{equation}
-		\bvec{z} \sim \mathcal{N}(1, 0)
+            F_A = \int_{\lambda_1}^{\lambda_2}\phi_A(\lambda)\left[xI_\lambda(T_h)+(1-x)I_\lambda(T_c)\right]d\lambda
         \end{equation}
-	then $\bvec{x} = \bvec{\mu_x} + L\bvec{z}$ is a vector with entries correlated by $K$ and with mean $\bvec{\mu_x}$, i.e.
+        where $\phi_A(\lambda)$ is the profile of the filter and $x$ is the covering fraction of the region at $T_h$.
+        As the covering fraction varies, the flux changes according to
         \begin{equation}
-		\bvec{x} \sim \mathcal{N}(\bvec{\mu_x}, K)
+            \frac{dF_A}{dx} = \int_{\lambda_1}^{\lambda_2}\phi_A(\lambda)\left[I_\lambda(T_h)-I_\lambda(T_c)\right]d\lambda
         \end{equation}
-	\gordontodo{fix this so that $\mu$ depends on $\theta$ and is a function of $\bvec{t}$ or something like that. Compuational cost should be $\mathcal{O}(NJM^2)$}
+        From this we see that the rate at which the flux varies with variations in $x$ depends on the 
+        contrast in brightness between the two regions within the filter. We therefore expect the flux 
+        to be more variable where the contrast between the two regions is higher than where it is lower. For 
+        instance, in the infrared, we should see higher amplitude variability features in filters 
+        centered at shorter wavelengths than longer ones.
 
-	\subsection{interpolation and extrapolation}	
-	\gordontodo{Bit more complicated --- if both dimensions are celerite processes then the scaling is better than in the case where $R$ is not celerite. You have notes on this.}
-
-\section{example: exomoons}
-	\gordontodo{What if I used Kepler-1625 as an example? Could simulate WFC3 GRISM data and show that moon could be validated with Celerite2D?}
+        We can expand the flux as a function of covering 
+        fraction in a Taylor series to first order: 
+        \begin{equation}
+            F_A(x) = F_A(\bar{x}) + \frac{dF_A}{dx}(x-\bar{x})
+        \end{equation}
+        %Putting this in terms of the deviation of the flux from $x=x_0$ we have
+        %\begin{equation}
+        %    \delta F_A = F_A(x)-F_A(x_0) = \frac{dF_A}{dx}(x-x_0)
+        %\end{equation}
+        where $\bar{x}$ is the mean covering fraction. We can then compute the covariance between the flux in filter $A$ at time $t_n$ and the flux in some filter $B$ at time $t_m$/ 
+        Filter $B$ is assumed to be centered at a longer wavelength than $A$:
+        \begin{eqnarray}
+            \nonumber \mathrm{cov}(F_A(t_n), F_B(t_m)) &=& \mathrm{E}\left[(F_A(t_n) - \bar{F_A})(F_B(t_m)-\bar{F_B})\right] \\
+            \nonumber &=& \frac{dF_A}{dx}\frac{dF_B}{dx}\mathrm{E}\left[(x_n-\bar{x})(x_m-\bar{x})\right] \\
+            &=& C_AC_B\mathrm{cov}(x_n, x_m)
+        \end{eqnarray}
+        where $C_A$ and $C_B$ are the derivatives of $F_A$ and $F_B$ with respect to 
+        covering fraction and we've used the fact that $\mathrm{cov}(x_n, x_m) = \mathrm{E}\left[(x_n-\bar{x})(x_m-\bar{x})\right]$. 
+        This result satisfies equation (\ref{eqn:multiplication_condition}) with $r(u_n, u_m) = C_nC_m$. The 
+        corresponding covariance matrix for the second dimension is 
+        \begin{equation}
+            [R]_{nm} = C_nC_m
+        \end{equation}
+    \subsection{Inferring transit parameters}
 
@@
 \appendix
+\section{The \celerite Algorithm}
+\label{sec:celerite_algorithm}
+\section{The Hadamard and Kronecker Products}
+\label{sec:matrix_prod}
+\section{Proof of \ref{eqn:KUVR}}
+\label{sec:Kproof}
+\section{Computing the Two-dimensional GP}
+\label{sec:algorithms}
 
@@
 
-\begin{thebibliography}{}
+%\begin{thebibliography}{celerite2d.bib}
 
@@
 
-\end{thebibliography}
+%\end{thebibliography}
+\bibliography{main.bib}
 
